---
title: Maps
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(maps)
library(mapproj)
library(ggthemes)
```

# Create Original Map

```{r}
states <- map_data("state")
ggplot(states, aes(x=long, y=lat, group=group)) + 
    geom_polygon(color="gold2", fill="navyblue")
```

## Your turn

Edit the code so that each state is a different color

```{r}
states <- map_data("state")
ggplot(states, aes(x=long, y=lat, group=group)) + 
    geom_polygon(color="gold2", fill="navyblue")
```

# Your turn: ACS data

Your task is to use the American Community Survey data to make a chloropleth map of the US

You should:

-   Use the starter code provided below
-   Choose a different variable
-   Change the color scale
-   Update the title, axis labels, and legend

## Load Data

```{r}
acs_state_data = read_csv("acs_state_data_2022_5y.csv") 
acs_state_data$state = tolower(acs_state_data$NAME)
```

Variable information:

-   `med_age`: median age
-   `med_income`: median age
-   `race_X`: estimated population of (self-reported) race
-   `born_in_state`: estimated population who were born in state
-   `X_age_married`: average age at first marriage for self-reported male and female respondents
-   `hs_diploma`, `associate_degree`, `prof_degree`, etc.: estimated number with a high school diploma, associate's degree, professional degree, etc.
-   `internet_X`: estimated number of people who have internet services

## ACS Data Starter Map

**Important note:** we're using `geom_map` as a shortcut here to make a chloropleth map. This saves us the step of encoding the lat/long borders of states, and lets us directly fill the state colors using `acs_state_data` *only*. If you're asked to make a chloropleth map with data that only includes, e.g., state name, use this starter code.

```{r}
ggplot(data = acs_state_data) + 
  geom_map(
    aes(map_id = state, fill = total_pop), 
    map = states
  ) +
  expand_limits(x = states$long, y = states$lat) +
  coord_map() +
  theme_map()
```
