---
title: "`bind_` and `_join`: Combining Data"
subtitle: Solutions
author: Stat220 -- W25
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE)
library(tidyverse)
library(ggthemes)
library(here)
```

Click the "code" button above to copy and paste the source code, or pull a fresh version of the ["activities" repo](https://github.com/stat220-s25/activities) from GitHub.

# Bakeoff Ratings (again)

We'll start with our `messy_ratings2` dataset from last week:

```{r}
messy_ratings2 <- read_csv("https://stat220-s25.github.io/data/messy_ratings2.csv")
messy_ratings2
```



```{r}
tidy_ratings2 <- messy_ratings2 %>%
  select(series, contains("7day")) %>%
  pivot_longer(-series, names_to = "episode", values_to = "viewers") %>%
  separate(episode, into = c("episode", "period")) %>%
  mutate(
    episode = parse_number(episode),
    period = parse_number(period)
  )
```


```{r}
tidy_ratings2 %>%
  ggplot(aes(x = episode, y = viewers, col = series, group = series)) +
  geom_line() + 
  scale_color_viridis_c(option = "plasma", end = .8)
```

## Task 

```{r}
messy_ratings2b <- read_csv(here(here(), "data/messy_ratings2_9_14.csv"))
```

### Combine with `bind_rows()`

```{r}
messy_ratings2 %>%
  bind_rows(messy_ratings2b)
```

### Combine with `join()`

First, create `tidy_ratings2b`

```{r}
tidy_ratings2b <- messy_ratings2b %>%
  select(series, contains("7day")) %>%
  pivot_longer(-series, names_to = "episode", values_to = "viewers") %>%
  separate(episode, into = c("episode", "period")) %>%
  mutate(
    episode = parse_number(episode),
    period = parse_number(period)
  )
```

Why doesn't `left_join()` work? 

```{r}
tidy_ratings2 %>%
  left_join(tidy_ratings2b)
```

Which `_join` do we use? 

```{r}
tidy_ratings2 %>%
  full_join(tidy_ratings2b)
```

### Combine with `bind_cols`


Why doesn't this work?

```{r}
seasoncols_ratings2 = tidy_ratings2 %>%
  pivot_wider(names_from = series, 
              values_from = viewers, 
              names_prefix = "s")

seasoncols_ratings2b = tidy_ratings2b %>%
  pivot_wider(names_from = series, values_from = viewers)

seasoncols_ratings2 %>%
  bind_cols(seasoncols_ratings2b)

```

Solve it by removing the episode/period

```{r}
seasoncols_ratings2b = tidy_ratings2b %>%
  pivot_wider(names_from = series, 
              values_from = viewers, 
              names_prefix = "s") %>%
  select(-c(episode, period))

seasoncols_ratings2 %>%
  bind_cols(seasoncols_ratings2b)

```

**Stop here** and let Amanda know you've finished

# `join` the episodes data

```{r}
tidy_ratings = tidy_ratings2 %>%
  full_join(tidy_ratings2b)
```

```{r}
episodes = read_csv("https://stat220-s25.github.io/data/bakeoff-episodes.csv")
```

```{r}
episodes %>%
  left_join(tidy_ratings, by = c("series", "episode")) %>%
  slice_max(viewers, n = 10, with_ties = TRUE) %>%
  select(signature)
```

Print the "signature" desserts that were made on the 10 episodes that had the highest 7-day viewership. 

**Stop here** and let Amanda know you've finished

# Pt 2: map data

join datasets and geom polygon
