---
title: "Data import and dates/times -- Solutions"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE,
                      eval = FALSE)
library(tidyverse)
library(ggthemes)
```

Click the "code" button above to copy and paste the source code

# Warm Up

Use `read_csv()` to import the `desserts` data set from

- `https://stat220-s25.github.io/data/desserts.csv`

```{r}
desserts <- read_csv("https://stat220-s25.github.io/data/desserts.csv")
```

# Your turn

Use the appropriate `read_<type>()` function to import the following data sets:

-   `https://stat220-s25.github.io/data/data-4.csv`

-   `https:/stat220-s25.github.io/data/tricky-1.csv`



If you hit any errors/problems, be sure to explore them and identify the issue, even if you can't "fix" it.

### data-4.csv

Opening the URL for `data-4.csv` you see that the delimiter is `|`, so you should use `read_delim()`:

```{r}
data4 <- read_delim("https://stat220-s25.github.io/data/data-4.csv", delim = "|")
```

### tricky-1.csv

At first glance it isn't obvious what is tricky about `tricky-1.csv`; however, if you try to use `read_csv()` you'll hit the following problem:

```{r}
tricky1 <- read_csv("https://stat220-s25.github.io/data/tricky-1.csv")
problems(tricky1)
```

The output of `problems()` here is that in rows 4 and 7 `read_csv` only finds 4 columns, but 5 are expected. Looking at the imported data set we see that the city is missing in rows three and 6 (the 4th and 7th rows of the original data file).

```{r}
tricky1
```

We can fix this with post processing:

```{r}
tricky1[3, ] <- c(tricky1[3, 1:2], NA, tricky1[3, 3:4])
tricky1[6, ] <- c(tricky1[4, 1], NA, tricky1[4, 3:5])
```

### tricky-2.csv

-   `https://stat220-s25.github.io/data/tricky-2.csv`

To begin, we can attempt to read in the `tricky-2.csv` file:

```{r}
tricky2 <- read_csv("https://stat220-s25.github.io/data/tricky-2.csv")
problems(tricky2)
```

This looks like a missing value problem again! Let's look at the rows with missing values:

```{r}

tricky2 %>% slice(8:21)
```

Why are there state abbreviations in the latitude column?!? (Because not all the rows put quotes around the "City, State" so it added an extra column!)

One solution is to parse the file in pieces. First, we can import the first 7 rows of the file:

```{r}
#| eval: false
tricky2_part1 <- read_csv(
  "https://stat220-s25.github.io/data/tricky-2.csv",
  n_max = 7
)
```

Next, we parse the other rows, and manually specify the column names, since we are skipping the first row:

```{r}
#| eval: false

tricky2_part2 <- read_csv(
  "https://stat220-s25.github.io/data/tricky-2.csv",
  skip = 8, 
  col_names = c("iata", "airport", "city", "state",  "latitude", "longitude")
)
```

Now, we have to use some data wrangling skills to combine everything. First, we must create a `state` column in `tricky2_part1`

```{r}
#| eval: false

tricky2_part1 <- tricky2_part1 %>%
  separate(city, into = c("city", "state"), sep = ",")
```

Then, we need to bind the two sets of rows together using `bind_rows()`:

```{r}
#| eval: false

tricky2 <- bind_rows(tricky2_part1, tricky2_part2)
```

# `read_excel` practice

Use the appropriate `read_<type>()` function to import the following data sets:

- `https://stat220-s25.github.io/data/sales.xlsx`

Step 0: `read_excel` doesn't like URLs, so you'll first have to visit this link in a browser, download the .xlsx file to your computer, and save it in the same folder as your activity .Rmd file

Step 1: read in the data so it looks like the following:

```
# A tibble: 9 × 2
  id      n    
  <chr>   <chr>
1 Brand 1 n    
2 1234    8    
3 8721    2    
4 1822    3    
5 Brand 2 n    
6 3333    1    
7 2156    3    
8 3987    6    
9 3216    5   
```

```{r}
library(readxl)
sales<- read_excel(here(here(), "data/sales.xlsx"),
           skip = 4,
           col_names = c("id", "n"))
```

Step 2 (Stretch goal): Manipulate the data so that it looks like the following:

```
# A tibble: 7 × 3
  brand   id    n    
  <chr>   <chr> <chr>
1 Brand 1 1234  8    
2 Brand 1 8721  2    
3 Brand 1 1822  3    
4 Brand 2 3333  1    
5 Brand 2 2156  3    
6 Brand 2 3987  6    
7 Brand 2 3216  5  
```

```{r}
sales |>
  mutate(
    brand = ifelse(n == "n", id, NA)
  ) |>
  fill(brand) |>
  filter(n != "n") |>
  select(brand, id, n)
```


# lubridate practice

```{r}
desserts2 <- read_csv("https://stat220-s25.github.io/data/desserts.csv")

desserts2 |>
  mutate(
    uk_airdate = dmy(uk_airdate),
    how_long_ago = today() - uk_airdate,
    how_long_ago = time_length(how_long_ago, unit = "year")
  ) |>
  select(
    series, episode, uk_airdate, how_long_ago
  )
```