---
title: "19-apis"
---

# TidyCensus

Install `tidycensus` if not on maize:

```{r}
#| eval: false
install.packages("tidycensus")
```

```{r}
library(tidycensus)
library(tidyverse)
```

## Example call:

```{r}
#| eval: false
acs_mn_2020 <- tidycensus::get_acs(
    year = 2020,
    state = "MN",
    geography = "county",
    variables = c("B01003_001", "B19013_001"),
    output = "wide",
    geometry = TRUE
)
```

## Storing your API key

1. Create a new text file in the same folder as your .rmd
2. Copy and paste your census key into the empty file
3. Save the file as `census_api_key.txt`

```{r}
#| eval: false
my_key <- readLines("census_api_key.txt")
```

and tell `tidycensus` what your API key is with:

```{r}
#| eval: false
census_api_key(my_key)
```

*Do not commit and push `census_api_key.txt` to github*

```{r}
#| eval: false
acs_mn_2020 |>
  mutate(
    name = str_remove(NAME, ", Minnesota"),
    name = str_remove(name, " County")
  ) |> 
  ggplot(aes(x = B19013_001E, 
                          xmin = B19013_001E - B19013_001M, 
                          xmax = B19013_001E + B19013_001M, 
                          y = fct_reorder(name, B19013_001E))) + 
  geom_point() + 
  geom_errorbarh() + 
  labs(
    x = "",
    y = ""
  )
```



## Links to more info about census

<https://data.census.gov/>

<https://www.census.gov/acs/www/data/data-tables-and-tools/subject-tables/>

## Try it

Add "educational attainment" to the topic filters and take a look at the first table. What is it showing you?

## Your turn

- Search for the "Median earnings in the past 12 months by major" table that we looked at on the census website using `load_variables`
- Run another call to `tidycensus::get_acs` to access this variable

```{r}
acs_vars <- load_variables(2020, "acs5", cache = TRUE) 
acs_vars 
```

# httr2

Install if not already installed:

```{r}
#| eval: false

install.packages("httr2")
```

```{r}
library(httr2)
```

## Example call

```{r}
#| eval: false
hmong_state_request <- request("https://api.census.gov/data") %>% 
    req_url_path_append("2019") %>% 
    req_url_path_append("acs") %>% 
    req_url_path_append("acs1") %>% 
    req_url_query(get = c("NAME", "B02015_009E", "B02015_009M"), `for` = I("state:*"), key = my_key, .multi = "comma")
```

```{r}
#| eval: false
hmong_state_response <- req_perform(hmong_state_request)
```

```{r}
#| eval: false
hmong_state_tbl <- hmong_state_response %>%
  resp_body_json(simplifyVector = TRUE) %>%
  janitor::row_to_names(1) %>%
  as_tibble()

hmong_state_tbl
```


## Your Turn

- Edit the `httr` code to access the `B15013_002E` variable from 2024 (median income for "computers, math, statistics" majors)
  - To get US level estimates, use `I("us:1")` in the `for` argument
  - To check your answer, copy the URL into a browser and make sure you get `"102338"`
- Make a `httr` request to access the 1-year ACS data from 2022 and 2023. Make sure to save your results from each call! 
- Combine all years into a single dataset (you may need to create a column for `year`)
- Make a time series plot with your chosen variable on the `y-axis` and year on the `x-axis`
    - You will need to do some cleaning of the data
- Brainstorm some ways to make the graph more interesting through gathering more data from the census API, and implement those ideas if you have time!

