---
title: "25-databases"
format: html
execute:
  eval: false
---

```{r}
library(tidyverse)
library(mdsr)
```

# Working with a remote database

```{r}
#| eval: false
db <- dbConnect_scidb("airlines")
flights <- tbl(db, "flights")
carriers <- tbl(db, "carriers")
airports <- tbl(db, "airports")
```

```{r}
#| eval: false

flights %>%
  head(5)
```


1. Of all destinations from MSP, which were the 10 most common in 2015? 

```{r}
#| eval: false

flights %>%
  filter(origin == "MSP", year == 2015) %>%
  group_by(dest) %>%
  summarize(
    n = n()
  ) %>%
  arrange(desc(n)) %>%
  head(10)
```

2. Which airport had the highest average arrival delay time in 2015? What about the lowest? 

```{r}
#| eval: false

flights %>%
  group_by(dest) %>%
  summarize(
    avg_arr_delay = mean(arr_delay, na.rm = TRUE)
  ) %>%
  arrange(desc(avg_arr_delay)) %>%
  head(10)
```

3. How many flights were there in 2015 between MSP and FAI (Fairbanks International Airport?) 

```{r}
#| eval: false

flights %>%
  filter(year == 2015) %>%
  filter((dest == "FAI" & origin == "MSP") | (origin == "FAI" & dest == "MSP")) %>%
  summarize(
    n = n()
  )
```

4. List the airline and flight number for all flights between MSP and FAI in September 2015

```{r}
#| eval: false

flights %>%
  filter(year == 2015, month == 9) %>%
  filter((dest == "FAI" & origin == "MSP") | (origin == "FAI" & dest == "MSP")) %>%
  select(carrier, flight)
```

# Practice with SQL commands

Repeat 1-4, but writing SQL commands

```{r}
#| eval: false

dbGetQuery(db, "SELECT faa, name FROM airports LIMIT 0,5")
```
1. Of all destinations from MSP, which were the 10 most common in 2015? 

```{r}
#| eval: false

dbGetQuery(db, 
"SELECT dest, SUM(1) AS numFlights
FROM flights 
WHERE origin = 'MSP' AND year = 2015
GROUP BY dest
ORDER BY numFlights DESC
LIMIT 0,10")
```

2. Which airport had the highest average arrival delay time in 2015? What about the lowest? 

```{r}
#| eval: false

dbGetQuery(db, 
"SELECT dest, AVG(arr_delay) AS avg_arr_delay
FROM flights 
WHERE year = 2015
GROUP BY dest
ORDER BY avg_arr_delay DESC
LIMIT 0,10")

dbGetQuery(db, 
"SELECT dest, AVG(arr_delay) AS avg_arr_delay
FROM flights 
WHERE year = 2015
GROUP BY dest
ORDER BY avg_arr_delay 
LIMIT 0,10")
```

3. How many flights were there in 2015 between MSP and FAI (Fairbanks International Airport?) 

```{r}
#| eval: false

dbGetQuery(db, 
"SELECT COUNT(*) AS N
FROM flights 
WHERE (dest = 'MSP' AND origin = 'FAI') OR (dest = 'FAI' AND origin = 'MSP')
")
```

4. List the airline and flight number for all flights between MSP and FAI in September 2015

```{r}
#| eval: false

dbGetQuery(db, 
"SELECT carrier, flight 
FROM flights
WHERE year = 2015 AND month = 9 AND ((dest = 'MSP' AND origin = 'FAI') OR (dest = 'FAI' AND origin = 'MSP'))
")

```

## Explain/test the following and translate to {dplyr} code


1. 
```
SELECT 
  year, month, day, origin, dest,
  flight, carrier
FROM flights
WHERE year = 2013 AND month = 6 AND day = 26
AND origin = 'BDL'
LIMIT 0, 6;
```

2. 

```
SELECT 
  name AS airport_name, 
  CONCAT('(', lat, ', ', lon, ')') AS coords
FROM airports
LIMIT 0, 6;
```

3. 

```
SELECT
  DISTINCT STR_TO_DATE(CONCAT(year, '-', month, '-', day), '%Y-%m-%d') 
    AS theDate
FROM flights
WHERE year = 2013 AND month = 6 AND day BETWEEN 26 and 30
  AND origin = 'BDL'
LIMIT 0, 6;
```
