---
title: Makeup Lab Quiz Info
author: Stat220 -- W25
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(tidyverse)
library(ggthemes)
library(nycflights23)
```

There will be a **make-up lab quiz** offered during Week 10. You can take this in-class makeup and replace your lowest in-class lab quiz (provided your score on the makeup is higher). There will be no take-home portion. 

 - Must schedule a 30 minute block outside of class during the following times: 
    - Monday 12:30-2:30pm
    - Tuesday 10:30a-12p
    - Wednesday 9-11am
    

# Guidelines

This is a *closed note*, *closed internet resources*, *closed other people* lab quiz. I want to see what's in your brain! You may use the cheat sheets provided by me, but otherwise you may not use any resources.

The lab quizzes are not written to be tricky or very difficult. If you've been completing the in-class activities and the homework, and putting the time and effort in to understand them, you should do well on the lab quizzes.

# Format

The reason I give lab quizzes is because your life will be easier if you know how to do basic tasks in R "on the fly". I also want to see what you **know** and not just what you can do with access to your resources. 

This makeup will only include a required in-class portion:


  - Pencil-and-paper questions
  - Cheat sheets provided by me but no other resources
  - Designed to assess what you know "on the fly" without access to resources


# Skills

**All skills from the first 3 lab quizzes are "fair game" for this makeup**, along with the following new topics. Since this quiz can replace a previous one, I'm looking for growth and proficiency with *all* topics from the course. 

## APIs and Web Scraping

- Know and explain the difference between using an API and web scraping
- Explain *simple* HTML scraping code

## Shiny

TBD

## SQL

TBD

# Grading

For the in-class portion, each question is worth 1 point, graded as "successful" (almost entirely correct), "not successful" (missing a key component), or "half credit" (mostly correct). 

# Practice Questions

## Data

The `nycflights23` package contains information about all flights that departed from NYC (e.g. EWR, JFK and LGA) in 2023. The main data is in the `flights` data frame, but there are additional data sets which may help understand what causes delays, specifically:

-   `weather`: hourly meteorological data for each airport
-   `planes`: construction information about each plane
-   `airports`: airport names and locations
-   `airlines`: translation between two letter carrier codes and names


```{r}
library(tidyverse)
library(nycflights23)
q <- 0
```

## Questions

TBD