---
title: Lab Quiz 01 Info
execute: 
  echo: true
  message: false
  warning: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE)
library(tidyverse)
library(ggthemes)
library(nycflights23)
```

Our first lab quiz is scheduled for **Friday of Week 3**. The first half of class will cover new content, and the second half of class you will complete the in-person portion of the lab quiz.

# Guidelines

This is a *closed note*, *closed internet resources*, *closed other people* lab quiz. I want to see what's in your brain! You may use the cheat sheets provided by me, but otherwise you may not use any resources.

The lab quizzes are not written to be tricky or hard. If you've been completing the in-class activities and the homework, and putting the time and effort in to understand them, you should do well on the lab quizzes.

# Skills

## R Basics

-   Given a vector, list, or data frame, extract an element of interest.
    -   Know how to use each extractor (`x$__`, `x[__]`, `x[[__]]`) and what they return
    -   Construct logical vectors to use as an index to subset a vector or data frame
    -   Construct integer vectors to use as an index to subset a vector or data frame
-   Given a vector, list, or data frame, obtain a quick summary such as the length, dimension, or type of each element
    -   Functions to know: `length()`, `nrow()`, `ncol()`, `dim()`, `summary()`, `glimpse()`, `class()`, `typeof()`, `head()`, `tail()`.

## Data Visualization

-   Identify the appropriate layer to add to a static graphic in order to display specific information
    -   You should know the `geom` (and associated aesthetics) for the following charts: bar/column chart, histograms, boxplots, density plots, violin plots, scatterplots, time series line plots, map, chloropleth map
-   Given a graph and a dataset structure, provide ggplot2 code to replicate the graphic:
    -   base layers
    -   axis labels, titles, captions
    -   `scale_x` functions for aesthetics
    -   `facet_wrap()` and `facet_grid()`
-   Given a question of interest and data set, decide on an appropriate graphic to address/answer the question
    -   You should know when the graphs mentioned above are appropriate
-   Given a graphic, describe the strengths and weaknesses of it from a design perspective
    -   Core principles:
    -   Accessibility considerations: color, alt text, direct labeling, etc.

## Data Wrangling (covered in Week 3)

-   Know how the following verbs act on a data set: `filter`, `slice`, `select`
    -   Syntax for using
    -   Describe what the output would look like
-   Given a data set and goal, identify and utilize the appropriate verb to create the data set of interest
-   Use {dplyr} verbs AND the logical vector approach to subset a data frame 

# Grading

For the in-class portion, each question is worth 1 point, graded as "successful" (almost entirely correct), "not successful" (missing a key component), or "half credit" (mostly correct). 

## Revision

::: callout-caution
Coming soon! 
:::


# Practice Quiz

## Data

The `nycflights23` package contains information about all flights that departed from NYC (e.g. EWR, JFK and LGA) in 2023. The main data is in the `flights` data frame, but there are additional data sets which may help understand what causes delays, specifically:

-   `weather`: hourly meteorological data for each airport
-   `planes`: construction information about each plane
-   `airports`: airport names and locations
-   `airlines`: translation between two letter carrier codes and names

For this lab quiz, we're going to work with a random sample of this data (called `flights_sampl`)

```{r}
library(tidyverse)
library(nycflights23)
q <- 0
```

## Example Questions

::: callout-caution
Coming soon! 
:::
