---
title: Makeup Lab Quiz Info
author: Stat220 -- W25
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(tidyverse)
library(ggthemes)
library(nycflights23)
```

There will be a **make-up lab quiz** offered during Week 10. You can take this in-class makeup and replace your lowest in-class lab quiz (provided your score on the makeup is higher). There will be no take-home portion. 

 Must schedule a 30 minute block outside of class during the following times: 
 
  - Monday 12:30-2:30pm
  - Tuesday 10:30a-12p
  - Tuesday 3-4pm
  - Wednesday 9-11am
    

# Guidelines

This is a *closed note*, *closed internet resources*, *closed other people* lab quiz. I want to see what's in your brain! You may use the cheat sheets provided by me, but otherwise you may not use any resources.

The lab quizzes are not written to be tricky or very difficult. If you've been completing the in-class activities and the homework, and putting the time and effort in to understand them, you should do well on the lab quizzes.

# Format

The reason I give lab quizzes is because your life will be easier if you know how to do basic tasks in R "on the fly". I also want to see what you **know** and not just what you can do with access to your resources. 

This makeup will only include a required in-class portion:


  - Pencil-and-paper questions
  - Cheat sheets provided by me but no other resources
  - Designed to assess what you know "on the fly" without access to resources


# Skills

**All skills from the first 3 lab quizzes are "fair game" for this makeup**, along with the following new topics. Since this quiz can replace a previous one, I'm looking for growth and proficiency with *all* topics from the course. 

## APIs and Web Scraping

- Know and explain the difference between using an API and web scraping
- Explain *simple* HTML scraping code

## Shiny

- Distinguish which code belongs in the `ui()` function and which belongs in the `server()` function
- Correctly track how an `input$id` flows from the UI into a reactive/rendering function in the server, and how it connects back to an `output$id` in the UI
- Know the matching pairs for displaying content (eg `plotOutput` matches `renderPlot`)

## SQL

- Given a SQL query, explain what output will be produced
- Translate between SQL queries and {dplyr} code

# Grading

For the in-class portion, each question is worth 1 point, graded as "successful" (almost entirely correct), "not successful" (missing a key component), or "half credit" (mostly correct). 

# Practice Questions

## Data

The `nycflights23` package contains information about all flights that departed from NYC (e.g. EWR, JFK and LGA) in 2023. The main data is in the `flights` data frame, but there are additional data sets which may help understand what causes delays, specifically:

-   `weather`: hourly meteorological data for each airport
-   `planes`: construction information about each plane
-   `airports`: airport names and locations
-   `airlines`: translation between two letter carrier codes and names


```{r}
library(tidyverse)
library(nycflights23)
q <- 0
```

## Questions

**Know and explain the difference between using an API and web scraping**


You are tasked with collecting daily temperature data for NYC airports. 

(a) Provide one reason why accessing an official weather API would be preferable to scraping a weather website's HTML
(b) If the weather website does *not* have an API

**Explain *simple* HTML scraping code**

Read the following {rvest} code and explain 

(a) What `html_elements("table.wikitable")` does
(b) What `html_table()` does

```{r}
#| eval: false


library(rvest)

url <- "https://en.wikipedia.org/wiki/List_of_busiest_airports_by_passenger_traffic"

busiest_airports <- read_html(url) |>
  html_elements("table.wikitable") |>
  html_table() |>
  pluck(1)
```

**Shiny**

Below is a very simple Shiny app. I've noted 3 errors related to reactivity and app structure. For each error, identify what is wrong and how to fix it. 

(a) Error 1: 

(b) Error 2: 

(c) Error 3: 

```{r}
#| eval: false 


ui <- fluidPage(
  selectInput("airport_choice", "Choose an Airport:", choices = c("JFK", "LGA", "EWR")),
  renderPlot("delay_plot") # Error 1
)

server <- function(input, output, session) {
  
  output$delay_plot <- plotOutput({ # Error 2
    
    plot_data <- flights |> 
      filter(origin == airport_choice) # Error 3
      
    ggplot(plot_data, aes(x = dep_delay)) + 
      geom_histogram()
  })
}

shinyApp(ui, server)
```

**Given a SQL query, explain what output will be produced**

Below is a SQL query using the `flights` and `airlines` tables in the SQL databse. (you can assume all tables/datasets and variable names are the same as {nycflights23}). Describe the output in 1-2 sentences. Your answer should include what each row and each column represent.

```
SELECT airlines.name, COUNT(flights.flight) AS total_flights
FROM flights
INNER JOIN airlines 
  ON flights.carrier = airlines.carrier
WHERE flights.origin = 'JFK'
GROUP BY airlines.name
ORDER BY total_flights DESC
LIMIT 5;
```


**Translate between SQL queries and {dplyr} code**

Below is a SQL query: 

```
SELECT origin, AVG(dep_delay) AS avg_delay
FROM flights
WHERE month = 1
GROUP BY origin
ORDER BY avg_delay DESC;
```

Write down a {dplyr} pipeline to accomplish the same task (you can assume all tables/datasets and variable names are the same)



